---
title: "Summary document"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}
 library(haven)
bee <- read_dta("Data/bee data new 14-15 v.5.25.dta")


summary()

```


Sunflower Honey Bee Heath

honeybee colony loss

neonicotinoids on honeybee disease and mortality

1.question 

whether/how timing of sunflower flowering differs between states 

entomological 


### Data


outcome variable :  Varroa mites and Nosema parasite loads as indicators of honey bee health  




variables : forage availability and weather.



crop layer, honey bee passing, correlation 

We spatially join the USDA Animal and Plant Health Inspection Services (APHIS) Survey of Honey Bee
Pests and Disease with NASS Cropscape data using the geographic coordinates of the apiaries and year
the sample was taken. We specifically focus on those colonies that are not migratory to ensure that the nearby cropscape appropriately represents the landscape at the time the sample was taken (Holt, 2014).

 


The USDA APHIS conducts the Honey Bee Pest and Disease Survey as a means of identifying pests,
pathogens, and disease affecting honey bees in the United States. This data set contains information on apiary samples collected from 2009 to 2014 throughout the United States. Forty states with 2552 samples are in the data set. 

In each sampled apiary, at least 8 colonies are tested for a number of diseases and pests. Not all samples are tested for pollen residue; only 676 samples have pollen sample results. Since there is no crop information for samples in Hawaii, we exclude these areas from our analysis as well.


Along with excluding migratory colonies, this leaves us with a sample of 429 apiaries in 33 states with most represented in the North East and the Mid-west (see Table 1)




Using geocoded data, we plot the raster density maps for Varroa mites and Nosema parasites for non-migratory apiaries with pollen results and find a correlation between the detection of neonics and higher levels of diseases (see Figure 1).


This suggests that bees’ exposure to neonics may potentially contribute to higher disease levels.

To estimate the degree of neonicotinoid exposure, we first identify the major field crops that are
traditionally seed treated with neonicotinoids, including corn, soy, cotton, canola, sorghum, barley, rice and wheat. 

Then we map the sampled non-migratory apiaries in APHIS onto NASS cropscape data
determine the crops grown within a 2-mile radius of each apiary. The resolution of these data is set at 30 meters squared per pixel (USDA NASS n.d.). We extract the crop area within two miles of each apiary as this is vicinity in which bees typically do most of their foraging (Eckert, 1933).

Therefore, this two mile area, which comprises over 8,000 acres, provides the best estimate of the crops and landscape that bees would interact with during their foraging. We then calculate the percentage of the two mile buffer area


occupied by each crop with the assumption that a linear relationship exists between changes in treated  crop area and morbidity loads.


The top land cover near all sampled apiaries is forested upland; it is comprised of deciduous forests,
mixed forests and evergreen forests and makes up about 25% of the 2-mile buffers on average. Several
neonicotinoid-treated crops are also among the top 10 landcover categories, including corn, soybean and
winter wheat (see Table 2). When we look at the land covers by contamination status of the apiaries, we
find that corn, apples and oranges are significantly more prevalent near apiaries where neonicotinoids
have been detected (see Table 3).
With geographic coordinates for non-migratory apiaries, we also extract potential forage and weather data
within the 2-mile buffers from NASS’ Vegscape layers and Oregon State University’s Prism database.
USDA NASS provides data about on the vegetation cover of the United States over the period from 2000-
2014. For this study, we use data on the Normalized Difference Vegetation Index (NDVI), which
measures the density of vegetation within pixels representing 15 acres of landscape (Mueller and
Minchenkov 2013). Data on the NDVI is provided on a daily, weekly and biweekly basis. We use apiaries’ locations as well as sample collection time, and obtain the biweekly NDVI data, which provides
information about the average NDVI within a 16 day window.



To control for the timing of exposure to pesticides and more abundant nectar sources, we collect
information on the time of planting and blooming for neonicotinoid-treated crops. NASS collects
agricultural plant timing data for select crops in some states. We compile information on the planting
percentage by month for the United States each year. Corn, soy, cotton, canola, rice, sorghum, barley and
spring wheat are planted in the spring. Winter wheat is planted in the fall. Most spring planting occurs
between April and June. Fall planting occurs between September and November. Due to a lack of
information, we estimate the planting window for canola to be from April 20 to June 10 every year
(Canola Council of Canada, 2013). Information on bloom timing of honeybee forage plants is provided on
HoneyBeeNet, which not only lists the plants that bees frequently forage within each region within each
state, but also whether each plant is significant nectar source or not (Nickeson, 2010).


Many of the honey bee forage crops fall into the natural area landscape category, so we consider the entire
area to be in bloom if at least one of the forage crops is in bloom within the natural area category. Bloom
timing for neonic-treated crops are also included in the data set to estimate pesticide exposure from
pollen. Spring wheat, winter wheat, barley and rice are not considered as forage crops for honey bees and
thus are missing bloom timing information from HoneyBeeNet. These crops are wind-pollinated are not
adapted to attract pollinators. However, honey bees can consume pollen from these crops (Burlew 2013).

To control for unobserved regional effects, we use USDA census regions. These regions are selected to
increase comparability with studies on overwintering losses and to isolate regional cropping patterns.
Forested upland, developed areas and grasslands are among the top 5 crops in every region (see Table 4).
Both corn and soybeans are very prevalent in USDA Region 1, which is includes most Northeastern states
and parts of the eastern Midwest including New York, New Jersey, Wisconsin, Ohio, Michigan, Illinois


This region largely coincides with the US corn belt and has 10 apiaries detected with neonicotinoids
among the total 18 contaminated ones. USDA Region 3 is made up of southern states, including Florida,
Louisiana, and Texas, and 5 apiaries in this region have been detected with neonicotinoids. The other 3
contaminated apiaries come from Regions 2, 4 and 6.




### Methods 

* "Do Neonics matter"

compare colonies located near neonicotinoid-
treated crops to those further away, during planting, blooming versus other times of year. 



take into account the biology, honey bee educators, honey bee patterns


For our analysis, we first use a logit regression to ask what factors are associated with finding
neonicotinoid contamination in the hive. Second, we use a multivariate regression to estimate the effect
of neonicotinoid contamination on colony Nosema and Varroa loads. We use several specifications, first
with no fixed effects, then adding fixed effects for region and year. Then we include other controls for
forage availability and weather. In the first stage, we compare those apiaries that are near neonicotinoid-
treated crops whose samples are taken during planting to other apiaries near neonocotinoid-treated crops
whose samples are taken other times of year, and to apiaries who are not near treated crops. For the
second analysis on health outcomes, using the fixed effects, we compare disease outcomes of those
apiaries where neonicotinoids are found to apiaries tested in the same region, in the same year and during
the same time of year.





### Results 

* Crop cover data, focus on sunflower acreage 

2-mile buffer 


The first stage examines whether an apiary with a large share of neonicotinoid-treated crops within the
foraging radius have a higher probability to being contaminated by neonicotinoids during certain times of
the year. In other words, we ask at which time of the year do we observe neonicotinoid exposure in the
apiary, and does this timing align with planting or blooming of neonicotinoid-treated crops. We aggregate
the percent area of all 9 commonly neonicotinoid-treated crops and interact these numbers with planting
and bloom time. Planting time and bloom time are both dummy variables indicating whether any treated
crops within the 2-mile radius are being planted or in bloom on the date of the sample collection




Crop acre*Bloom time
Crop acreage
Planting time
Crop acre*Planting time
Bloom time

Minimum Temperature
NDVI
Precipitatio




In the second stage, we examine whether being contaminated by neonicotinoids is associated with an
increase in disease loads.

We run the same regressions with mites loads as the outcome variable and find little evidence that
neonicotinoid contamination affects Varroa mite levels

Nosema Parasite /  Varroa Mites
Neonics
NDVI
Minimum Temperature
Precipitation



take a paragraph to say where we have sunflowers and where we don;t

divide by state, in our sample, 


seasonality, sun flower blooming, control for the timing, in 2km radius, control for the timing and control for location





#### summary statistics 

how many observations have any sunflower within the 2 mi radius 



when were those samples were taken


* take the lat and lon of the sampleid that we already have in the data 



extract the acreage information from the NASS Cropscape data?


##### correlation 


geocode bee hives on crop model, diversity of colonies of sun flower

##### regression 


summary stats on the existing, comparison within the sate (near)


it seems almost 90% of the sampleid have 0 acreage of sunflowers around them.

11%, from 900 up to 2418300, I guess the unit is km squared.
sorry, that's 85 observations


summary statistics, correlation   





oct sept, sampling, try for may-june lited time frames (late summer and fall)



 
create a simple summary table that says number of observations by month and year, and by state for obs with and w/o sunflower?  (you can drop the neonic stuff too)

group 0 is no sunflower, group 1 is at least 900m^2 sunflower acreage within the 2km radius.


